Evolution of pre-copulatory and post-copulatory inbreeding behaviour and associated polyandry

### 2016-11-01 ms_pre_post.Rmd
Remove some of the empirical/theory language from the manuscript; this really needs to be written as an empirical paper, and hence one result of many across taxa

Inbreeding avoidance has been widely hypothesised to evolve through pre-copulatory and post-copulatory mechanisms as a consequence of inbreeding depression in offspring viability. However, no modelling compares evolution of pre-copulatory and post-copulatory inbreeding avoidance jointly, and given biologically relevant costs of inbreeding strategies. Critically, evolution of post-copulatory inbreeding avoidance is restricted to females that mate multiply, potentially leading to complex evolutionary dynamics among inbreeding strategies and polyandry. Accordingly, we use individual-based modelling to track evolution of polyandry and inbreeding strategies under multiple costs and constraints to address three questions: Does evolution of post-copulatory inbreeding avoidance facilitate evolution of costly polyandry? Is evolution of one costly inbreeding strategy inhibited by another non-costly alternative strategy? And is evolution of one inbreeding strategy precluded by existence of another already adaptive strategy. We show that evolution of polyandry is facilitated by post-copulatory inbreeding avoidance. Evolution of relatively low cost inbreeding avoidance strategies was observed, while selection for functionally redundant high cost strategies decreased over generations. Finally, existence of an adaptive inbreeding strategy precluded evolution of post-copulatory, but not pre-copulatory, inbreeding avoidance. Our model thereby introduces a novel framework for predicting evolution of both pre-copulatory and post-copulatory inbreeding strategies in empirical systems.

@@

Inbreeding often greatly reduces the fitness of inbred offspring [termed 'inbreeding depression'; @Charlesworth1999; @Keller2002; @Charlesworth2009]. Such strong inbreeding depression is widely hypothesised to drive evolution of inbreeding avoidance, which can be enacted through multiple forms of mate choice in species with obligate biparental reproduction [@Parker1979; @Parker2006; @Pusey1996; @Szulkin2012]. From the perspective of a focal breeding female, inbreeding avoidance might be achieved by avoiding mating with related males (i.e., pre-copulatory inbreeding avoidance), or by biasing fertilisation towards unrelated males after mating has occurred, meaning that related males are less likely to sire offspring (i.e., post-copulatory inbreeding avoidance). Importantly, evolution of post-copulatory inbreeding avoidance requires that females are polyandrous, defined as mating with multiple males during a single reproductive bout. Further, such polyandry could itself evolve because it allows females to mate with additional unrelated males following initial mating with a relative, thereby facilitating pre-copulatory inbreeding avoidance even in the absence of any post-copulatory female choice or otherwise biased fertilisation (e.g., given a 'fair raffle' among sperm provided by feamles' mates). As a reproductive strategy, polyandry can thereby simultaneously allow females to mate with less closely related males and creates the opportunity for further inbreeding avoidance enacted through active post-copulatory choice. Polyandry has consequently been widely hypothesised to evolve as an adaptation to allow females to avoid inbreeding and thereby increase offspring fitness [@Zeh1997; @Jennions2000; @Tregenza2002; @Akcay2007]. However, despite such well-established hypotheses, no modelling investigates conditions under which pre-copulatory or post-copulatory inbreeding avoidance, or both strategies, are likely to evolve, or predicts how such evolution will feed back to affect underlying evolution of polyandry.

@@

Meanwhile, empirical studies of diverse species have found evidence for female inbreeding avoidance enacted through pre-copulatory or post-copulatory routes, or both. For example, pre-copulatory inbreeding avoidance has been documented in weevils, birds, and butterflies [e.g., @Kuriwada2011; @Kingma2013; @Fischer2015], and post-copulatory inbreeding avoidance has been documented in crickets [e.g., @Simmons2006; @Bretman2009]. Evidence of both pre-copulatory and post-copulatory inbreeding avoidance has accumulated for some well-studied species, including Trinidadian guppies [*Poecilia reticulata*; @Gasparini2011; @Daniel2015] and house mice [*Mus domesticus*; @Potts1991; @Firman2015]. While most studies consider either pre-copulatoryor post-copulatory inbreeding avoidance in isolation, a small number of studies consider both. Accordingly, @Liu2014 found evidence of pre-copulatory, but not post-copulatory, inbreeding avoidance in cabbage beetles (*Colaphellus bowringi*). Other studies found no evidence of pre-copulatoryor post-copulatory inbreeding avoidane despite strong inbreeding depression [e.g., @Reid2014; @Reid2015]. These empirical studies demonstrate that diverse combinations of pre-copulatoryand post-copulatory inbreeding avoidance occur in nature. However, there is as yet no modelling framework that predicts *a priori* what combinations should be favoured by selection when both pre-copulatory and post-copulatory inbreeding avoidance, and polyandry can evolve. Such a framework would allow prediction of the form of mating system evolution across different systems that experience inbreeding depression and allow existing empirical studies to be compared to model predictions.

@@

A recent genetically-explicit individual-based model showed that selection for alleles underlying polyandry and pre-copulatory inbreeding avoidance, and resulting joint evolution of these reproductive strategies, occurred only under highly constrained conditions [@Duthie]. Specifically, such evolution occurred when direct costs of polyandry were small, when very fewe males were available for a female's initial mate choice but many were available for additional mate choice(s), or when polyandry was conditionally expressed when a focal female was related to her initial mate [@Duthie]. In the absence of these conditions, increasing degrees of polyandry tended to increase the overall degree of inbreeding that females experienced, and hence increase inbreeding depression in offspring fitness. This occurred because polyandrous females that had evolved to avoid inbreeding had already chosen available unrelated males as their initial mates. Their additional mates, chosen from the remaining male population, were therefore increasingly likely to include relatives. Evolution of polyandry purely to facilitate pre-copulatory inbreeding avoidance was consequently restricted.

However, while @Duthie considered simultaneous evolution of polyandry and pre-copulatory inbreeding avoidance, they did not also allow simultaneous evolution of post-copulatory inbreeding avoidance, and assumed that polyandrous females had no post-copulatory control over fertilisation. If post-copulatory inbreeding avoidance could also evolve, polyandrous females could further bias paternity towards unrelated males within their set of chosen mates. This process might reduce the cost of polyandry stemming from accumulation of related mates, potentially facilitating evolution of polyandry to avoid inbreeding [@Zeh1997], and driving further consequent evolution of pre-copulatory or post-copulatory mate choice strategies. Therefore, while pre-copulatory inbreeding avoidance can facilitate evolution of polyandry under some restrictive conditions [@Duthie], the opportunity for post-copulatory choice might facilitate evolution of polyandry under a much wider range of conditions by allowing females to bias fertilisation toward non-kin. The long-term evolutionary dynamics of pre-copulatoryand post-copulatory inbreeding avoidance then become unclear. Strong inbreeding depression might drive initial evolution of both pre-copulatoryand post-copulatory inbreeding avoidance and associated polyandry. However, if evolution of polyandry and post-copulatory inbreeding avoidance renders pre-copulatory inbreeding avoidance redundant, or vice versa, then only one strategy might be maintained in the long-term.

Previous models implicitly or explicitly considered the fate of a rare allele underlying pre-copulatory inbreeding avoidance in a population initially fixed for random mating [e.g., @Parker1979; @Parker2006; @Duthie; @Duthie2016a]. Such models thereby isolate the invasion fitness of this single strategy. However, when both pre-copulatoryand post-copulatory strategies can affect realised inbreeding, it cannot be assumed that both strategies will invade simultaneously, nor that invasion fitness of one strategy will be independent of the other For example, if pre-adaptation or rapid selection results in the fixation of alleles underlying adaptive pre-copulatory inbreeding avoidance, then alleles underlying post-copulatory inbreeding avoidance might be precluded from invading a population in which post-copulatory choice is not already expressed because the effect of such invading alleles could be weak or negligible.

Ultimately, the absolute and relative frequencies of alleles underlying pre-copulatoryand post-copulatory inbreeding avoidance and polyandry are likely to be affected by direct negative selection on resulting phenotypes (i.e., costs) and hence on the net costs versus benefits stemming from inbreeding avoidance. Polyandrous females can pay costs of mate searching or mating, for example, because these activities increase predation risk [e.g., @Rowe1988; @Ronkainen1994; @Koga1998]. Females that express pre-copulatory choice (e.g., reluctance to mate with relatives) might pay a cost of increased risk of mating failure, or harm caused by sexual conflict [@Rowe1994; @Kokko2013]. Females that express post-copulatory choice might pay an energetic cost associated with physiological or biochemical mechanisms needed to store sperm and successfully bias fertilisation [e.g., @Gasparini2011; @Tuni2013; @Fitzpatrick2014b]. If the relative costs affecting pre-copulatoryand post-copulatory inbreeding avoidance differ, then alleles underlying the less costly inbreeding strategy might become fixed in the population over generations, while alleles underlying the more costly inbreeding strategy might become extinct if their effects are both costly and made redundant by the less costly strategy.

Here we use individual-based modelling to set up *in silico* experiments, which test the hypothesis that, given strong inbreeding depression in offspring viability, evolution of post-copulatory inbreeding avoidance facilitates evolution of polyandry. Additionally, we test the hypothesis that costs associated with polyandry and inbreeding strategies will strongly affect evolution such that only the less costly inbreeding avoidance strategy will be expressed in the long term. Finally, we test the hypothesis that evolution of one inbreeding avoidance strategy (i.e., pre-copulatoryor post-copulatory) will be precluded if it is introduced into a population that is already fixed for the other inbreeding avoidance strategy that is adaptive. We thereby provide a framework for predicting evolution of pre-copulatory and post-copulatory inbreeding avoidance across different reproductive systems.

@@

We have used *in silico* experiments to test three key hypotheses concerning evolution of pre-copulatory and post-copulatory inbreeding avoidance and associated polyandry. The modelling that we develop provides a novel framework for approaching empirical studies of inbreeding avoidance, predicting evolution of polyandry and inbreeding avoidance through pre-copulatoryand post-copulatory mechanisms. Accordingly, we modelled evolution of three separate traits affecting polyandry and inbreeding avoidance, including tendency for polyandry, pre-copulatory inbreeding strategy, and post-copulatory inbreeding strategy.

<!--- Using *in silico* experiments, we tested three specific hypothesis to predict evolution of polyandry, and pre-copulatoryand post-copulatory inbreeding strategy. We found that (1) selection for polyandry was greatly strengthened given concurrent evolution of post-copulatory inbreeding avoidance, (2) selection for costly pre-copulatoryor post-copulatory inbreeding avoidance was negligible given evolution of a less costly strategy of inbreeding avoidance, and (3) pre-existence of adaptive pre-copulatory inbreeding avoidance precluded evolution of polyandry and post-copulatory inbreeding avoidance, but not vice versa. --->

The opportunity to adjust inbreeding has been widely considered as a driver of adaptive evolution of polyandry [@Tregenza2002; @Foerster2003; @Akcay2007; @Varian-Ramos2012; @Kingma2013; @Lehtonen2015; @Reid2014]. Recently, @Duthie found the conditions under which selection for polyandry and pre-copulatory inbreeding avoidance is predicted to be highly restrictive. However, @Duthie assumed that fertilisation of offspring was unbiased with respect to mate kinship, meaning that females could not express post-copulatory choice. In contrast, we found that when post-copulatory inbreeding avoidance could evolve, selection for polyandry was greatly strengthened (Figure 1). The proposition that polyandry might facilitate cryptic female choice among males of varying compatibility is not new [e.g., @Zeh1997], but the existence of selection for post-copulatory inbreeding avoidance causing feedback to select for polyandry has not been formally modelled. Our model therefore has widespread implications for empirical studies of evolution of polyandry.

@@

In our model, one cost free inbreeding strategy precluded another more costly inbreeding strategy from persisting in a focal population (Figure 2), meaning that long-term persistence of both pre-copulatoryand post-copulatory inbreeding avoidance might not be expected in populations given cost asymmetry. Direct costs therefore strongly affected the evolution of the costly phenotype(s), meaning that quantifying costs is necessary for predicting evolution of phenotypes associated with inbreeding strategies. Quantifying direct costs of phenotypes associated with polyandry and mate choice is empirically challenging [@Pomiankowski1987; @Kokko2003; @Reid2015]. @Pomiankowski1987 categorised four types of costs that are relevant to mate choice, which we suggest are also relevant costs for polyandry, and include costs of elevated risks associated with (1) predation or (2) disease transmission, and costs incurred through (3) loss of time or (4) depletion of energy. In our model, we interpreted a cost of polyandy as an elevated risk of predation as might be incurred while searching for or courting mates [e.g., @Rowe1988; @Rowe1994], a cost of pre-copulatory inbreeding strategy as a lost of time [i.e., risk of not finding a mate in time due to choosiness; e.g., @Kokko2013], and a cost of post-copulatory inbreeding strategy as depletion of energy. While these costs are biologically realistic, different costs will be relevant for different populations and be realised in different ways. For example, increased risk of disease transmission has been observed for highly polyandrous females [@Roberts2015a], but such a cost would more realistically be applied to the number of realised mates a female has rather than her tendency for engaging in polyandry. Future models will therefore need to carefully consider the magnitudes and realisation of direct costs to accurately predict evolution of inbreeding strategies in populations.

>>>>

Inbreeding avoidance has been widely hypothesised to evolve through pre-copulatory and post-copulatory mechanisms as a consequence of inbreeding depression in offspring viability. However, no modelling compares evolution of pre-copulatory and post-copulatory inbreeding avoidance jointly, and given biologically relevant costs of inbreeding strategies. Critically, evolution of post-copulatory inbreeding avoidance is restricted to females that mate multiply, potentially leading to complex evolutionary dynamics among inbreeding strategies and polyandry. Accordingly, we use individual-based modelling to track evolution of polyandry and inbreeding strategies under multiple costs and constraints to address three questions: (1) Does evolution of post-copulatory inbreeding avoidance facilitate evolution of costly polyandry? (2) Is evolution of one costly inbreeding strategy inhibited by another non-costly alternative strategy? And (3) is evolution of one inbreeding strategy precluded by existence of another already adaptive strategy? We show that evolution of polyandry is facilitated by post-copulatory inbreeding avoidance. Evolution of relatively low cost inbreeding avoidance strategies was observed, while selection for functionally redundant high cost strategies decreased over generations. Finally, existence of an adaptive inbreeding strategy precluded evolution of post-copulatory, but not pre-copulatory, inbreeding avoidance. Our model thereby introduces a novel framework for predicting evolution of both pre-copulatory and post-copulatory inbreeding strategies in biological systems.

@@

<!--- I have removed 'mate' in front of 'choice' and replaced it with 'reproductive decisions' in the second sentence because post-copulatory inbreeding avoidance isn't an expression of mate choice; mating has already happened at this point; females are choosing paternity among the mates that they already chose --->

Inbreeding often greatly reduces the fitness of inbred offspring [termed 'inbreeding depression'; @Charlesworth1999; @Keller2002; @Charlesworth2009]. Such strong inbreeding depression is widely hypothesised to drive evolution of inbreeding avoidance, which can be enacted through multiple reproductive decisions in species with obligate biparental reproduction [@Parker1979; @Parker2006; @Pusey1996; @Szulkin2012]. From the perspective of a focal breeding female, inbreeding avoidance might be achieved by avoiding mating with related males (i.e., pre-copulatory inbreeding avoidance), or by biasing fertilisation towards unrelated males after mating has occurred, meaning that related males are less likely to sire offspring (i.e., post-copulatory inbreeding avoidance). Importantly, evolution of post-copulatory inbreeding avoidance requires that females are polyandrous, defined as mating with multiple males during a single reproductive bout. Further, such polyandry could itself evolve because it allows females to mate with additional unrelated males following initial mating with a relative, thereby facilitating pre-copulatory inbreeding avoidance even in the absence of any post-copulatory female choice or otherwise biased fertilisation (e.g., given a 'fair raffle' among sperm provided by feamles' mates). As a reproductive strategy, polyandry can thereby simultaneously allow females to mate with less closely related males and creates the opportunity for further inbreeding avoidance enacted through active post-copulatory choice. Polyandry has consequently been widely hypothesised to evolve as an adaptation to allow females to avoid inbreeding and thereby increase offspring fitness [@Zeh1997; @Jennions2000; @Tregenza2002; @Akcay2007]. However, despite such well-established hypotheses, no modelling investigates conditions under which pre-copulatory or post-copulatory inbreeding avoidance, or both strategies, are likely to evolve, or predicts how such evolution will feed back to affect underlying evolution of polyandry.

@@

Meanwhile, evidence for female inbreeding avoidance enacted through pre-copulatory or post-copulatory routes, or both, has been found across diverse taxa. For example, pre-copulatory inbreeding avoidance has been documented in weevils, birds, and butterflies [e.g., @Kuriwada2011; @Kingma2013; @Fischer2015], and post-copulatory inbreeding avoidance has been documented in crickets [e.g., @Simmons2006; @Bretman2009]. Evidence of both pre-copulatory and post-copulatory inbreeding avoidance has accumulated for some well-studied species, including Trinidadian guppies [*Poecilia reticulata*; @Gasparini2011; @Daniel2015] and house mice [*Mus domesticus*; @Potts1991; @Firman2015]. While most studies consider either pre-copulatory or post-copulatory inbreeding avoidance in isolation, a small number of studies consider both. Accordingly, @Liu2014 found evidence of pre-copulatory, but not post-copulatory, inbreeding avoidance in cabbage beetles (*Colaphellus bowringi*). Other studies found no evidence of pre-copulatory or post-copulatory inbreeding avoidance despite strong inbreeding depression [e.g., @Reid2014; @Reid2015]. These studies demonstrate that diverse combinations of pre-copulatory and post-copulatory inbreeding avoidance occur in nature. However, there is as yet no modelling framework that predicts *a priori* what combinations should be favoured by selection when both pre-copulatory and post-copulatory inbreeding avoidance, and polyandry, can evolve. Such a framework would allow prediction of the form of mating system evolution across different systems that experience inbreeding depression and allow existing studies to be compared to model predictions.

@@

A recent genetically-explicit individual-based model showed that selection for alleles underlying polyandry and pre-copulatory inbreeding avoidance, and resulting joint evolution of these reproductive strategies, occurred only under highly constrained conditions [@Duthie]. Specifically, such evolution occurred when direct costs of polyandry were small, when very few males were available for a female's initial mate choice but many were available for additional mate choice(s), or when polyandry was conditionally expressed when a focal female was related to her initial mate [@Duthie]. In the absence of these conditions, increasing degrees of polyandry tended to increase the overall degree of inbreeding that females experienced, and hence increase inbreeding depression in offspring fitness. This occurred because polyandrous females that had evolved to avoid inbreeding had already chosen available unrelated males as their initial mates. Their additional mates, chosen from the remaining male population, were therefore increasingly likely to include relatives. Evolution of polyandry purely to facilitate pre-copulatory inbreeding avoidance was consequently restricted.

However, while @Duthie considered simultaneous evolution of polyandry and pre-copulatory inbreeding avoidance, they did not also allow simultaneous evolution of post-copulatory inbreeding avoidance, and assumed that polyandrous females had no post-copulatory control over fertilisation. If post-copulatory inbreeding avoidance could also evolve, polyandrous females could further bias paternity towards unrelated males within their set of chosen mates. This process might reduce the cost of polyandry stemming from accumulation of related mates, potentially facilitating evolution of polyandry to avoid inbreeding [@Zeh1997], and driving further consequent evolution of pre-copulatory or post-copulatory mate choice strategies. Therefore, while pre-copulatory inbreeding avoidance can facilitate evolution of polyandry under some restrictive conditions [@Duthie], the opportunity for post-copulatory choice might facilitate evolution of polyandry under a much wider range of conditions by allowing females to bias fertilisation toward non-kin. The long-term evolutionary dynamics of pre-copulatory and post-copulatory inbreeding avoidance then become unclear. Strong inbreeding depression might drive initial evolution of both pre-copulatory and post-copulatory inbreeding avoidance and associated polyandry. However, if evolution of polyandry and post-copulatory inbreeding avoidance renders pre-copulatory inbreeding avoidance redundant, or vice versa, then only one strategy might be maintained in the long-term.

Previous models implicitly or explicitly considered the fate of a rare allele underlying pre-copulatory inbreeding avoidance in a population initially fixed for random mating [e.g., @Parker1979; @Parker2006; @Duthie; @Duthie2016a]. Such models thereby isolate the invasion fitness of this single strategy. However, when both pre-copulatory and post-copulatory strategies can affect realised inbreeding, it cannot be assumed that both strategies will invade simultaneously, nor that invasion fitness of one strategy will be independent of the other For example, if pre-adaptation or rapid selection results in the fixation of alleles underlying adaptive pre-copulatory inbreeding avoidance, then alleles underlying post-copulatory inbreeding avoidance might be precluded from invading a population in which post-copulatory choice is not already expressed because the effect of such invading alleles could be weak or negligible.

Ultimately, the absolute and relative frequencies of alleles underlying pre-copulatory and post-copulatory inbreeding avoidance and polyandry are likely to be affected by direct negative selection on resulting phenotypes (i.e., costs) and hence on the net costs versus benefits stemming from inbreeding avoidance. Polyandrous females can pay costs of mate searching or mating, for example, because these activities increase predation risk [e.g., @Rowe1988; @Ronkainen1994; @Koga1998]. Females that express pre-copulatory choice (e.g., reluctance to mate with relatives) might pay a cost of increased risk of mating failure, or harm caused by sexual conflict [@Rowe1994; @Kokko2013]. Females that express post-copulatory choice might pay an energetic cost associated with physiological or biochemical mechanisms needed to store sperm and successfully bias fertilisation [e.g., @Gasparini2011; @Tuni2013; @Fitzpatrick2014b]. If the relative costs affecting pre-copulatory and post-copulatory inbreeding avoidance differ, then alleles underlying the less costly inbreeding strategy might become fixed in the population over generations, while alleles underlying the more costly inbreeding strategy might become extinct if their effects are both costly and made redundant by the less costly strategy.

Here we use individual-based modelling to set up *in silico* experiments, which test the hypothesis that, given strong inbreeding depression in offspring viability, evolution of post-copulatory inbreeding avoidance facilitates evolution of polyandry. Additionally, we test the hypothesis that costs associated with polyandry and inbreeding strategies will strongly affect evolution such that only the less costly inbreeding avoidance strategy will be expressed in the long term. Finally, we test the hypothesis that evolution of one inbreeding avoidance strategy (i.e., pre-copulatory or post-copulatory) will be precluded if it is introduced into a population that is already fixed for the other inbreeding avoidance strategy that is adaptive. We thereby provide a framework for predicting evolution of pre-copulatory and post-copulatory inbreeding avoidance across different reproductive systems.

@@

We have used *in silico* experiments to test three key hypotheses concerning evolution of pre-copulatory and post-copulatory inbreeding avoidance and associated polyandry. The modelling that we develop provides a novel framework for approaching questions concerning inbreeding avoidance, and for predicting evolution of polyandry and inbreeding avoidance through pre-copulatory and post-copulatory mechanisms. Accordingly, we modelled evolution of three separate traits affecting polyandry and inbreeding avoidance, including tendency for polyandry, pre-copulatory inbreeding strategy, and post-copulatory inbreeding strategy.

<!--- Using *in silico* experiments, we tested three specific hypothesis to predict evolution of polyandry, and pre-copulatory and post-copulatory inbreeding strategy. We found that (1) selection for polyandry was greatly strengthened given concurrent evolution of post-copulatory inbreeding avoidance, (2) selection for costly pre-copulatory or post-copulatory inbreeding avoidance was negligible given evolution of a less costly strategy of inbreeding avoidance, and (3) pre-existence of adaptive pre-copulatory inbreeding avoidance precluded evolution of polyandry and post-copulatory inbreeding avoidance, but not vice versa. --->

The opportunity to adjust inbreeding has been widely considered as a driver of adaptive evolution of polyandry [@Tregenza2002; @Foerster2003; @Akcay2007; @Varian-Ramos2012; @Kingma2013; @Lehtonen2015; @Reid2014]. Recently, @Duthie found the conditions under which selection for polyandry and pre-copulatory inbreeding avoidance is predicted to be highly restrictive. However, @Duthie assumed that fertilisation of offspring was unbiased with respect to mate kinship, meaning that females could not express post-copulatory choice. In contrast, we found that when post-copulatory inbreeding avoidance could evolve, selection for polyandry was greatly strengthened (Figure 1). The proposition that polyandry might facilitate cryptic female choice among males of varying compatibility is not new [e.g., @Zeh1997], but the existence of selection for post-copulatory inbreeding avoidance causing feedback to select for polyandry has not been formally modelled. Our model therefore has widespread implications for understanding evolution of polyandry.

@@

In our model, one cost free inbreeding strategy precluded another more costly inbreeding strategy from persisting in a focal population (Figure 2), meaning that long-term persistence of both pre-copulatory and post-copulatory inbreeding avoidance might not be expected in populations given cost asymmetry. Direct costs therefore strongly affected the evolution of the costly phenotype(s), meaning that quantifying costs is necessary for predicting evolution of phenotypes associated with inbreeding strategies. Quantifying direct costs of phenotypes associated with polyandry and mate choice is empirically challenging [@Pomiankowski1987; @Kokko2003; @Reid2015]. @Pomiankowski1987 categorised four types of costs that are relevant to mate choice, which we suggest are also relevant costs for polyandry, and include costs of elevated risks associated with (1) predation or (2) disease transmission, and costs incurred through (3) loss of time or (4) depletion of energy. In our model, we interpreted a cost of polyandy as an elevated risk of predation as might be incurred while searching for or courting mates [e.g., @Rowe1988; @Rowe1994], a cost of pre-copulatory inbreeding strategy as a lost of time [i.e., risk of not finding a mate in time due to choosiness; e.g., @Kokko2013], and a cost of post-copulatory inbreeding strategy as depletion of energy. While these costs are biologically realistic, different costs will be relevant for different populations and be realised in different ways. For example, increased risk of disease transmission has been observed for highly polyandrous females [@Roberts2015a], but such a cost would more realistically be applied to the number of realised mates a female has rather than her tendency for engaging in polyandry. Future models will therefore need to carefully consider the magnitudes and realisation of direct costs to accurately predict evolution of inbreeding strategies in populations.
